---
title: "Introduction to R - Group Project"
output: html_notebook
---

## *Author*: "Tatjana Nill and Omogbolahan Alli",

### Problem: Problem Statement: Find high attack Countries, traffic exchange and attack duration to prevent future attacks!

Step1: Identify high attacked countries
Step2: identify pattern of attack duration or exchange data
Step3: Are the attacked channels the same?
Step4: What is the duration of the attack time, are they really critical - boot force or just penetration test?
Step5: Conclusion on findings

This is an [Cybersecurity: Suspicious Web Threat Interactions](https://www.kaggle.com/datasets/jancsg/cybersecurity-suspicious-web-threat-interactions) Project.

First we need to prepare our environment by loading required libraries.
```{r LOAD: Environment}
library(tidyverse)
library(ggplot2)
library(shiny)
library(dplyr)

```

Load data from my project location.
```{r LOAD: DATA}
# Read the dataset (adjust the path accordingly)
CloudAttack <- read.csv("~/HULT/Introduction-R/Project/CloudWatch_Traffic_Web_Attack.csv")
#Beggining of data Massaging & feature engineering
is.Date(CloudAttack$creation_time) #Check if Creation Time is a Data
is.Date(CloudAttack$end_time) #Check if End Time is a Data
is.Date(CloudAttack$time) #Check if  Time is a Data


table(CloudAttack$time) # what insights can we draw from time
table(CloudAttack$creation_time) # what insights can we draw from start time of attack
table(CloudAttack$end_time) # what insights can we draw from end time of attack


```

Step1: Identify high attacked countries
Count the attacks per Country
```{r COUNT: Attack per Country}

# Step 1: Identify high-attacked countries
high_attacked_countries <- CloudAttack %>%
  group_by(src_ip_country_code) %>%
  summarise(Attacks = n()) %>%
  arrange(desc(Attacks))

# Create a bar chart for attacked countries
ggplot(high_attacked_countries, aes(x = reorder(src_ip_country_code, Attacks), y = Attacks)) +
  geom_bar(stat = "identity", fill = "#1b98e0") +
  labs(title = "Number of Attacks by Country",
       x = "Country",
       y = "Number of Attacks") +
  theme_minimal()  # Customize the plot theme as desired

# Find duration of each attack
CloudAttack$attack_duration <- CloudAttack$end_time - CloudAttack$creation_time
View(CloudAttack)
table(CloudAttack$attack_duration) # All the attacks are 10 minutes long
table(CloudAttack$src_ip) # Looks like the attacks are coming from a handful of IPs
# Looks like the data isn't clean. 
# We have US showing up alongside NL, IL(Illinios), CA (California)
table(CloudAttack$src_ip_country_code) 

```
Step2: identify pattern of attack duration or exchange data

```{r ANALYSIS: indentify pattern}
# Show Ip adresses that attacks the US 



# Show source IP country 
```

Which IPs should the US blacklist to reduce the amount of attacks

```{r RESULT: Outcome}




`````````



