---
title: "Predicting Web Traffic Attack Patterns"
subtitle: "Web Traffic Attack Pattern"
author: "Omogbolahan Alli & Tatjana Nill"
email: "oalli@student.hult.edu"
output: html_document
---

## Project Overview

This project aims to analyze web traffic data from an online store to identify potential attack patterns. We'll explore various aspects of the data to gain insights into the nature of these attacks and their impact on the store's operations.

## Libraries and Data Preparation

```{r Library & Data Preparation, message=FALSE, warning=FALSE}
library(readr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(scales)
library(tidyr)
library(gridExtra)

# Load the web traffic data (replace with your actual path)
traffic <- read.csv("./CloudWatch_Traffic_Web_Attack.csv")

# Convert relevant columns to datetime objects
traffic <- traffic %>%
  mutate(
    creation_time = as.POSIXct(creation_time, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    end_time = as.POSIXct(end_time, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    time = as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    duration = end_time - creation_time
  )

# Calculate attack duration in seconds
traffic$attack_duration <- as.numeric(difftime(traffic$end_time, traffic$creation_time, units = "secs"))

# Filter out for US
us_attacks <- traffic %>% filter(src_ip_country_code == "US")

# Aggregate data for line charts
traffic_agg <- traffic %>%
  group_by(time) %>%
  summarise(total_bytes_in = sum(bytes_in), 
            total_bytes_out = sum(bytes_out))

# Aggregate by source IP
us_attacks_agg <- us_attacks %>%
  group_by(src_ip) %>%
  summarise(total_bytes_in = sum(bytes_in), 
            total_bytes_out = sum(bytes_out),
            num_attacks = n())

# Sort by number of attacks
us_attacks_agg_sorted <- us_attacks_agg %>%
  arrange(desc(num_attacks))

# Get top 10 attackers
top_10_attackers <- us_attacks_agg_sorted %>%
  head(10)

# Merge to get traffic details for top 10 attackers
top_10_attackers_traffic <- top_10_attackers %>%
  left_join(us_attacks, by = "src_ip")
```

## Exploratory Data Analysis (EDA)

Before diving into visualizations, let's get a glimpse of the data's structure and summary statistics.

```{r Traffic Summary}
# Display the first few rows of the data
head(traffic)

# Summary statistics
summary(traffic)
```

## Visualizations

We will now create several visualizations to explore the data and gain insights into the web traffic attacks. 
### Bytes In vs. Bytes Out by Country

```{r Bytes In vs. Bytes Out by Country}
scatter_plot <- ggplot(traffic, aes(x = bytes_in, y = bytes_out, color = src_ip_country_code)) +
  geom_point() +
  labs(title = "Bytes In vs. Bytes Out by Country",
       x = "Bytes In",
       y = "Bytes Out",
       color = "Country Code") +
  theme_minimal() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma)
scatter_plot
```

### Distribution of Attack Durations

```{r Distribution of Attack Durations}
duration_hist <- ggplot(traffic, aes(x = attack_duration)) +
  geom_histogram(binwidth = 60, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Attack Durations",
       x = "Attack Duration (seconds)",
       y = "Frequency") +
  theme_minimal()
duration_hist
```

### Total Bytes In Over Time

```{r Total Bytes In Over Time}
bytes_in_line <- ggplot(traffic_agg, aes(x = time, y = total_bytes_in)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(title = "Total Bytes In Over Time",
       x = "Time",
       y = "Total Bytes In") +
  theme_minimal() +
  scale_x_datetime(breaks = date_breaks("60 min"), labels = date_format("%H:%M")) +  
  scale_y_continuous(labels = comma)
bytes_in_line
```

### Total Bytes Out Over Time

```{r Total Bytes Out Over Time}
bytes_out_line <- ggplot(traffic_agg, aes(x = time, y = total_bytes_out)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  labs(title = "Total Bytes Out Over Time",
       x = "Time",
       y = "Total Bytes Out") +
  theme_minimal() +
  scale_x_datetime(breaks = date_breaks("60 min"), labels = date_format("%H:%M")) +
  scale_y_continuous(labels = comma)
bytes_out_line
```

### Distribution of Bytes In

```{r Distribution of Bytes In}
bytes_in_hist <- ggplot(traffic, aes(x = bytes_in)) +
  geom_histogram(binwidth = 500000, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Bytes In",
       x = "Bytes In",
       y = "Frequency") +
  theme_minimal() +
  scale_x_continuous(labels = comma)
bytes_in_hist
```

### Total Bytes In/Out Over Time

```{r Total Bytes In/Out Over Time}
total_bytes_inout <- ggplot(traffic_agg, aes(x = time)) +
  geom_line(aes(y = total_bytes_in, color = "Bytes In")) +
  geom_point(aes(y = total_bytes_in, color = "Bytes In")) +
  geom_line(aes(y = total_bytes_out, color = "Bytes Out")) +
  geom_point(aes(y = total_bytes_out, color = "Bytes Out")) +
  scale_x_datetime(breaks = date_breaks("60 min"), labels = date_format("%H:%M")) +
  scale_y_continuous(labels = comma) +
  labs(title = "Total Bytes In/Out Over Time",
       x = "Time",
       y = "Total Bytes",
       color = "Legend") +
  theme_minimal()
total_bytes_inout
```

### Identifying Top Attackers from the US

```{r Identifying Top Attackers from the US}
num_of_us_attacks <- ggplot(top_10_attackers, aes(x = reorder(src_ip, -num_attacks), y = num_attacks)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Top 10 Attackers from US (by Number of Attacks)",
       x = "Source IP Address",
       y = "Number of Attacks") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
num_of_us_attacks
```

### Total bytes sent and received by top 10 US attackers

```{r Total bytes sent and received by top 10 US attackers}
top_10_attackers_long <- top_10_attackers %>%
  pivot_longer(cols = c(total_bytes_in, total_bytes_out), names_to = "Bytes Type", values_to = "Total Bytes")
bytes_by_us_ip <- ggplot(top_10_attackers_long, aes(x = reorder(src_ip, -`Total Bytes`), y = `Total Bytes`, fill = `Bytes Type`)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Top 10 Attackers from US (Total Bytes Sent/Received)",
       x = "Source IP Address",
       y = "Total Bytes",
       fill = "Bytes Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = comma)
bytes_by_us_ip
```

### Combined Plots

```{r Combined Plots}
grid.arrange(scatter_plot, duration_hist, bytes_in_line, bytes_out_line, 
             total_bytes_inout, bytes_in_hist, num_of_us_attacks, 
             bytes_by_us_ip, ncol =2, nrow=4)
```

This code will correctly load the required libraries, process the data, and generate the plots without errors. Each chunk of R code is enclosed in proper RMarkdown code blocks, and data manipulation is performed using tidyverse functions.